\section{\underline{One Dimensional Harmonic Oscillator}}

1. \underline{Review}

First let us review the standard treatment of the linear harmonic oscillator in the coordinate representation which you can find in practically all text books of Quantum Mechanics, e.g, Pauling \& Wislon or Schiff pp. 67 ff. I will assume that you 
have already done this in a more elementary course in Quantum Mechanics and so I'll only review the more important aspects here. 

The harmonic oscillator or something closely related to it is forever appearing the physics of all types of phenomena -- mechanics, E \& M, solid state, etc. -- and is often used as the starting point to solving more complicated sytems. 

Our Hamiltonian is
$$H = {p^2\over 2m} + \half k x^2 = {p^2\over 2m}  + \half  m \omega^2 x^2  \mbox{   where  } \omega^2\equiv {k \over m}.$$
Since $H$H is time independent, we can write $\ds \Psi(x,t) = \psi(x)e^{-{iEt\over\hbar}}$ and our wave equation 
$$H\Psi(x,t) = i\hbar{\partial \over \partial t}\Psi(x,t)$$ reduces to 
$$H\psi(x) = E \psi(x) \mbox{   or  } \left( {\hbar^2\over 2m}{d^2\over dx^2} + E - \half m \omega^2 x^2 \right)\psi(x) = 0.$$
At this point, it is convenient to change variables and we define
$$E = \lambda\left({\hbar\omega\over 2}\right) \mbox{    and    } x = \xi\sqrt{ {\hbar\over m \omega}}.$$
Our differential equations becomes
$${d^2 \psi(\xi)\over d\xi^2} + (\lambda - \xi^2)\psi(\xi) = 0.$$
For large $\xi$ we can neglect $\lambda$ and we see that our solution goes like $\ds e^{-{\xi^2\over 2}}$. So we look for a solution of the form
$$\psi(\xi) = \phi(\xi)\exp^{-{\xi^2\over 2}}$$ which gives us a differential equation of 
$${d^2\phi\over d\xi^2} - 2\xi {d\phi\over d\xi} + (\lambda - 1)\phi = 0.$$
Now a standard approach is to try a series solution of the form
$$\psi(\xi) = \xi^s(b_o + b_1 \xi + b_2\xi^2 \hdots)\mbox{   where } b_0 \ne 0.$$
Substituting this solution into the differential equation we find by equating each coefficient of each power of $\xi$ to zero that $s = 0\mbox{  or } 1$ from the 1st equation and so $s=0 \mbox{ and/or } b_1 =0$ from the 2nd
equation. The succeeding equations five use recursion relations so that we have
\begin{eqnarray*}
\phi(\xi) &=& a_0 + a_2 \xi^2 + a_4 \xi^4 + \hdots \mbox{   for } s = 0  \mbox{ and}\\ 
\phi(\xi) &=& a_1\xi + a_3 \xi^3 + a_5 \xi^5 + \hdots \mbox{   for } s = 1  \mbox{ with}\\
a_{k+2} &=& {2k -\lambda + 1\over (k+1)(k+2)} a_k.
\end{eqnarray*}

One can quite easily show that both solutions $\ds \psi(\xi) = \exp^{-{lxi^2\over2}}\phi(xi)$ $\mbox{ for } s=0 \mbox{ or } 1$ diverge as $x\goes \pm \infty$.But this is not acceptable as $\psi(x)$ must remain finite.
$\psi(x)$ can only remain finite if $\phi(\xi)$ is a finite series which will occur if $\lambda$ is restricted to only take on the values $2n + 1$ where $n=0,1,2,\hdots$. This restricts the possible values of $E$ and we find
the quantized energy eigenvalues of $$E_n = (n + \half)\hbar\omega.$$
By picking the two arbitrary constants $a_0$ and $a_1$ ``correctly'', the $phi(\xi)$'s become the Hermite polynomials:
\begin{eqnarray*}
H_0(\xi) &=& 1\\
H_1(\xi) &=& 2\xi\\
H_2(\xi) &=& 4\xi^2 - 2i\\
H_3(\xi) &=& 8\xi^3 - 12\xi2i\\
\hdots &\null& \hdots\\
H_n(\xi) &=& (-1)^n \exp^{{\xi^2\over 2}}{d^n e^{-\xi^2}\over d\xi^n}.
\end{eqnarray*}
$H_n(\xi)$ satisfies the differential equation of $\phi(\xi)$ with $\lambda = 2n+1$, viz. 
$$H_n'' - 2\xi H'_n + 2n H_n = 0.$$
One can also show that 
\begin{eqnarray*}
H'_n &=& 2n H_{n-1}\\
0 &=& H_{n+1} - 2\xi H_n + 2n H_{n-1}\\
H_{n+1} &=& 2\xi H_n - H'_n
\end{eqnarray*}
In fact, one can take this last equation along with $H_0(\xi) = 1$ and ``generate'' all the other $H_n$'s. 

One can also use a generating function to obtain the $H_n$'s. The generating function $g(\xi,h)$ is given by
$$g(\xi,h) = e^{2\xi h - h^2} = \sum_{n=0}^\infty\, {H_n(\xi)h^n\over n!}.$$
The $H_n$'s are obtained by expanding $\ds e^{2\xi h - h^2}$ and equating coefficients of like powers of $h$ on both sides. These generating functions are vey useful in evaluating integrals involving the $H_n$'s. 

One usually normalizes $\psi(x)$ to unity by
$$\ket{n} \equiv \psi_n(\xi) = {1\over\sqrt{2^n n!\sqrt{\pi}}}e^{-{\xi^2\over 2}} H_n(\xi)$$ and then 
$$\int\, \psi^*_n(x) \psi_m(x)\, dx = \sqrt{{\hbar\over m\omega}} \int\, \psi^*_n(\xi) \psi_m(\xi)\, d\xi = \delta_{nm}\mbox{  and}$$
$$H\psi_n(\xi) = E_n\psi_n(\xi).$$
The matrix elements $\xAy{n}{H}{m}$ form a diagonal matrix in the coordinate representation
\[ H = \begin{pmatrix}
\half\hbar\omega & 0  & 0 & 0& 0& \hdots  \\
0  & {3\over2} \hbar\omega & 0&  0& 0& \hdots  \\
0 & 0   & {3\over2}\hbar\omega& 0& 0& \hdots \\
0 & 0   & 0 & {7\over2}\hbar\omega & 0& \hdots \\
\hdotsfor[2]{6}\\
\end{pmatrix} \]
and one can show that 
\[ \xi  = \begin{pmatrix}
0 & \sqrt{1\over2} & 0  & 0 & 0& 0& \hdots  \\
\sqrt{1\over2} & 0  & \sqrt{2\over2} & 0&  0& 0& \hdots  \\
0& \sqrt{2\over2}  & 0   & \sqrt{3\over2}& 0& 0& \hdots \\
0& 0 & \sqrt{3\over2}   & 0 & \sqrt{4\over2} & 0& \hdots \\
\hdotsfor[2]{7}\\
\end{pmatrix} \] \label{4-VII}

\subsection{\underline{2. An Alternatice Approach}}
After this quick review of the harmonic oscillator, I now want to show you what I promised earlier. That is, given the fundamental observables of the system and their commutation relations, I can 
solve this system without ever going into a representation. In fact, one should always be able to do this in principle but it is generally not easy to do so in practive. 

In this case, our fundamental observables are $p$ and $q$ (I will now use $q$ instead of $x$) and we have
$$\left[ p,p\right] = \left[q,q\right] = 0\mbox{   and  } \left[q,p\right] = i\hbar.$$
All our observables and/or operators are functions of $p$ and $q$. In particular
$$H = {p^2\over2m} + \half m\omega^2 q^2 = .$$
Instead of using $p$ and $q$, it is more conventient to introduce two other operators $R^\dagger$ and $R$ where 
\begin{eqnarray*}
R^\dagger &\equiv& {1\over\sqrt{2m\omega\hbar}}(m\omega q - ip)\mbox{~~~~~        and} \\
R               &\equiv& {1\over\sqrt{2m\omega\hbar}}(m\omega q + ip)
\end{eqnarray*}
or inversely
\begin{eqnarray*}
q &=& \sqrt{\hbar\over2m\omega} (R^\dagger + R)\mbox{~~~~~         and } \\
p &=& \sqrt{m\omega\hbar\over2} (R^\dagger  - R)
\end{eqnarray*}

Using $[q,p] = i\hbar$, we find that 
$$[R,R^\dagger] = 1, \quad [R^\dagger, R^\dagger] = 0, \mbox{~~~~~and } [R,R] = 0.$$
For $H$, we find 
$$H = (R^\dagger R + \half)\hbar\omega = (R^\dagger R + RR^\dagger){\hbar\omega\over2}.$$
So our eigenvalue equaton
$$H\ket{\Psi} = E\ket{\Psi}$$ reduces to solving an eigenvalue equation for
$R^\dagger R$. That is, if we can solve 
$$R^\dagger R \ket{n} = \lambda_n \ket{n}$$ then due to the form of $H$, we can certainly find $E_n$ in 
$$H\ket{n} = E_n \ket{n}.$$ 

First we show that $\lambda_n$ is non-negative. For this, consider $\ket{\chi} \equiv R\ket{n}$. Then 
$$\braket{\chi}{\chi} = \xAy{m}{R ^\dagger R}{n}  = \lambda_n\braket{n}{n}$$ and since $\braket{\chi}{\chi}$ and $\braket{n}{n}$ are positive definite, we need
$$\lambda_n \ge 0.$$

Next we look at
$$\ket{\chi} \equiv R^\dagger \ket{n}$$ and operate on it with $R^\dagger R$. Then
\begin{eqnarray*}
R^\dagger R\ket{\chi} &=& R^\dagger R R^\dagger\ket{n} = R^\dagger(1+ R^dagger R)\ket{n} = R^\dagger(1+\lambda_n)]ket{n}\\
                                    &=& (\lambda_n + 1)(R^\dagger\ket{n}) = (\lambda_n + 1) \ket{\chi}.
\end{eqnarray*}

So we see that $\ket{\chi} = R^\dagger \ket{n}$ is an eigenket of $R^\dagger R$ with an eigenvalue $(\lambda_n + 1)$. Likewise $R^\dagger R^\dagger \ket{n}$ is an eigenket with 
eigenvalue $(\lambda_n+2)$. Thus, given $\ket{n}$, we can ``generate'' all the higher eigenkets of $R^\dagger R$.

In a like manner, we can show that $R\ket{n}$ is also an eigenfunction of $R^\dagger R$ with an eigenvalue of $\lambda_n-1$...and so forth for $RR\ket{n}$, $RRR\ket{n}, \hdots$. 
While we could increase $\lambda_n$ indefinitely by using $R^\dagger$'s, we cannot reduce it indefinitely by using $R$'s as we must satisfy $\lambda_n \ge 0$ and so we cannot have
negative eigenvalues. If we operate on each successively smaller eigenket with $R$, we will eventually reach the eigenket with the ``lowest possible'' eigenvalue. 
Let's call this eigenket $\ket{\lambda_0}$. Then upon operating with $R$ again, we must have
$$R\ket{\lambda_0} = 0$$ or otherwise we will generate an eigenket with a lower eigenvalue which will be a contradiction. Now $\lambda_0$ is somewhere between zero and one. We have
$$R^\dagger R\ket{\lambda_0} = \lambda_0\ket{\lambda_0} \mbox{~~~~~ and } R\ket{\lambda_0} = 0.$$
But $R^\dagger R\ket{\lambda_0} = R^\dagger (R\ket{\lambda_0}) = R^\dagger (0) = 0$ which tells us that $\lambda_0 = 0$ uniquely.

Instead of labeling our eigenkets by $\ket{\lambda_0}, \ket{\lambda_1}, \ket{\lambda_2}, \hdots,$ let us use $\ket{0}, \ket{1}, \ket{2}, \hdots, \ket{n}, \hdots$. For these eigenkets, we have
$$R^\dagger R\ket{n} = n \ket{n}$$ with $n = 0,1,2,\hdots$ and $n$ has no upper limit. 

Clearly since $H= (R^\dagger R + \half)\hbar\omega$, we have
$$H\ket{n} = (n+\half) \hbar\omega = E_n\ket{n}.$$
At this point, we have found the eigenspectrum of the Hamiltonian $H$ which has eigenvalues of $\half\hbar\omega, {3\over2}\hbar\omega, {5\over2}\hbar\omega,\hdots$ and we did it without
going into an explicit representation -- That is, we are still working on our abstract Hilbert space. It is also clear that $H$ is diagonal in terms of these eigenkets. 

To complete the picture, all we have to do now is demonstrate how one obtains or calculates matrix elements of $p$ and $q$ or or arbitrary functions of $p$ and $q$.

First we make $R^\dagger \ket{n} \propto \ket{n+1}$ a normalized eigenket. Let $\ket{\chi}\equiv NR^\dagger\ket{n}$ and we want $\braket{\chi}{\chi} = 1$ along with 
$\braket{n}{n} = 1$. So we have
$$\braket{\chi}{\chi} = |N|^2\xAy{n}{RR^\dagger}{n} = |N|^2 \xAy{n}{1 + R^\dagger R}{n} = |N|^2(n+1)\braket{n}{n} = 1.$$
So we pick (up to an arbitrary phase) $\ds N = {1\over\sqrt{n+1}}$. This
$$\begin{matrix}
\mbox{Thus}     & \ket{n+1} &=& {1\over\sqrt{n+1}}R^\dagger \ket{n}\\ \\ 
\mbox{Likewise} &\ket{n-1} &=& {1\over\sqrt{n}}R\ket{n} 
\end{matrix}~~~~
\left\{\begin{matrix}
\mbox{~``Rule'' -- Use square  }\\ 
\mbox{~root of largest}\\
\mbox{~eigenvalue appearing} 
\end{matrix}\right. $$

Now we are ready to look at the matrix elements of $\ds \xi = q\sqrt{{m\omega\over\hbar}}$ -- that is, the matrix elements of $q$. 
As $\ds \xi = {1\over\sqrt{2}}(R^\dagger R)$, we see that 
$$\xAy{n}{\xi}{m} = {1\over\sqrt{2}}\xAy{n}{R^\dagger R}{m} = \sqrt{m+1\over2}\braket{n}{m+1} + \sqrt{m\over2}\braket{n}{m-1}$$ or 
$$\xAy{n}{\xi}{m} = \sqrt{m+1\over2}\delta_{n,m+1} + \sqrt{m\over2}\delta_{n, m-1}.$$ \underline{Really easy!!} Compare with $\xi$-matrix on page \cpageref{4-VII}. 
For the homework, we will evaluate $\xAy{n}{\xi}{m}$ using generating functions and you will appreciate how easy this method is.

Clearly we can continue and obtain the matrix elements of $q^2, p$, or any combinations of $p$'s and $q$'s using this approach.

Thus we have solved the harmonic oscillator without going into an explicit representation. If one wishes, he can start at this point and ``go into'' the coordinate representation. For details
see Schiff pages 184 and 185. 

\subsection{\underline{3. A Different Point of View}}
We now look at this new approach of using the $R^\dagger$'s and $R$'s in a different manner. Suppose we consider some elementary excitations which we call ``particles'' or ``quanta'' and 
which have an energy of $\hbar\omega$. Then we could say that the state $\ket{n+1}$ had one more of these ``particles'' in it than the state $\ket{n}$. Similarly the state $\ket{n-1}$ has one 
``particle'' less. Taking this point of view, $R^\dagger$ would when acting on a state create an additional one of these particles and $R$ would annihilate or destroy one of these particles. 
That is $R^\dagger$ and $R$ are creation and destruction operators for these particles respectively. In this interpretation $R^\dagger R$ is the counting operator which ``counts'' the number of 
paricles in a given state. 

These creation and destruction operators obey the commutation relation 
$$[R^\dagger. R^\dagger] = 0, \quad [R,R] = 0,\mbox{   and } [R,R^\dagger] = 1.$$
Particles which do this are called bosons and there wave functions are symmetric under interchange of any two particles. Furthermore, there is no restriction on how many one can
put into the same state with an energy $\hbar\omega$. 

To see this symmetry under interchange of any two particles, consider the state 
$$\ket{n} = {1\over\sqrt{n!}}R^\dagger R^\dagger \hdots R^\dagger\ket{0} = {1\over\sqrt{n!}}(R^\dagger)^n\ket{0}.$$
Perhaps is is clearer to put a subscript labelling each particle on the $R^\dagger$'s. Then 
$$\ket{n} = \ket{1 2 3 \hdots n} = {1\over\sqrt{n!}}R_n^\dagger R_{n-1}^\dagger\hdots R_3^\dagger R_2^\dagger R_1^\dagger \ket{0}.$$
Now we see that interchanging any two particles will give the same wave function back since the $R^\dagger$'s commute. 

An obvious generalization of the above approach is to have many linear oscillators all of different frequencies labeld by $\omega_k$ and with respective creation and annihilarion operators $R_k^\dagger$ and $R_k$.
Our commutation relations are 
$$[R_k, R_l] = 0, [R_k^\dagger, R_l^\dagger] = 0 \mbox{  ~~~ and  } [R_k, R_l^\dagger] = \delta_{kl}.$$

Lager on in the year we will be studying radiation from atoms. There we will see that the electromagnetic field can be represented by an infinite set of harmonic oscillators of all frequencies. We will then
``quantize'' these oscillators and the bosons resulting from this quantization will be the quanta of the electromagnetic field -- Photons. 

If our creation and annhilation operators had obeyed anticommutation relations instead of commutation relations, we would have a system of fermions instead of bosons. Then one can show that
the wave function is antisymmetric on interchange of any two particles and, at most, only one particle can occupy a given state at a time.

When one describes a system by using creation and annihiltion operators for the particles of the system, one says that he is using the ``occupation number representation'' or the ``formulation of second quantization''.

A glimpse of ``bigger and better'' things? If we can quantize the electromagnetic field, why can't we quantize the field describing the wave function $\psi(\vec{r})$? People have tried this and they 
originally hoped that the quanta of these fields would somehow be connected to elementary particles. But when people tried to quantize the field $\psi(\vec{r})$, they ran into all kinds of 
difficulties -- mostly of a mathematical nature, These problems have not been resolved and their treatment is beyond the scope of this course. 